---
title: "MA [46]15 - Data Science in R"
subtitle: "Covid plots from the class"
author: "Daniel Sussman"
date: "2022-02-04"
output: github_document
---



```{r}
knitr::opts_chunk$set(tidy = FALSE, cache = FALSE) # turn off formatR's tidy
knitr::opts_chunk$set(comment = "") # remove comment tags
knitr::opts_chunk$set(cache.path = "cache/", fig.path='figures/', fig.asp=.66, fig.width = 7, fig.align='center', dev="svg")
```


```{r}
library(tidyverse)
load("../lectures/data/covid19datahub/covid_updated_all_states.RData") 
load("../lectures/data/covid19datahub/covid_updated_big6.RData") 
```



---

# Coding in general

Also, I try to make the code be organized by separating lines, but error appears when running the code.
`Error: Cannot use `+.gg()` 

---

# Plots and ggplot questions

## Plots by date

```{r}
ggplot(covid) +
  geom_line(aes(x = date, y = tests, color = state))
```


I would like to make the data look less crowded and spread out the line that show on the plot.

---

```{r}
ggplot(covid_6, aes(x=date,y=tests/population, color = state)) + geom_point()+geom_col()
```


---

```{r}
ggplot(covid_6,aes(x=date,y=tests,color=state))+geom_area()
```

---

```{r}
ggplot(data=covid_6) + geom_line(aes(date, people_vaccinated, color = state))
```


---

```{r}
ggplot(weekly, aes(x = date, y = tests)) +
  geom_point(aes(color = as.factor(c2_workplace_closing)), size = .2) +
  geom_smooth(aes(linetype = state), size = 0.1)
```



---

```{r}
ggplot(data = covid_6) +
  geom_smooth(aes(x = date, y = deaths/confirmed, color = state)) +
  xlab("Date") +
  ylab("Death Rate (per case)") +
  labs(color = "State") +
  ggtitle("Covid Death Rate Over Time") +
  scale_y_continuous(labels=scales::percent)
```



---

```{r}
ggplot(weekly_6,aes(x=date,y=(confirmed/population)))+
  geom_line(aes(color=state),size=1.1)+
  labs(
    title = "Confirmed cases per capita in 6 states",
    subtitle = "Rolling weekly average from the start of 2020 to present",
    caption = "Graph created by Jacob Stenstrom for MA415"
  ) +
  ylab("Confirmed cases per capita") +
  xlab("Date")
```

Create a heat map with the darkest colored states being the one with the highest case count; reorder a column/bar graph to be in descending order.

---


## Adding facets

```{r fig.height=20, fig.width=10}
ggplot(data=covid)+
  geom_point(mapping=aes(x=date,y=confirmed/population,color=state))+
  facet_wrap(~state,nrow=8)
```

I wonder if there's a way to make the x-axis(dates) look cleaner so that it's not overlapping on each other and more readable.

---

```{r}
ggplot(covid, aes(x = date, y = deaths / confirmed)) +
  geom_point(aes(color = state), size = 0.1) +
  facet_wrap(~ as.factor(population))
```


---

```{r fig.height = 3, fig.width = 10}
library(RColorBrewer)
number_ticks <- function(n) {function(limits) pretty(limits, n)}

ggplot(data = covid_6)+
  geom_line( aes(x = date, y = tests/10000, color = state)) +
  scale_color_brewer(palette="Dark2")+
  ylab("every 10,000 tests")+
  facet_grid(~state)
```

I was unable to change the x scale to customize the date ticks.


---

```{r}
ggplot(covid_6) +
  geom_smooth(mapping = aes(x = date, y = deaths/population*100)) +
  labs(y = "Percentage Death per Population") +
  labs(title = "Deaths per Population Over Time") +
  facet_wrap(~state)
```

Would like to format the dates to look nicer.

---

## Coloring by a continuous variable

```{r}
ggplot(covid_6)+geom_point(aes(x=date, y=log(deaths), color = stringency_index), size=1, position = 'jitter')+
       geom_line(aes(x=date,y=log(people_fully_vaccinated), color=stringency_index), size=1.5)+
       facet_wrap(~state)+theme(axis.text.x=element_text(angle=30))+
       labs(title='(Log) Cumulative Deaths and Vaccinations Over Time at Varying Stringency Levels', x='Date Since Tracking Began', y='Cumulative Deaths (lower) and Vaccinations(upper) in log scale', color='Level of Stringency')
```

I still don't know how to label the lines without another package, but I may be missing something.
 
---

```{r}
ggplot(covid_6, aes(x = date, y = deaths, color = vaccines)) + geom_point() + facet_wrap(~ state) + ylab("deaths due to covid") + ggtitle("Covid Deaths")
```

---

```{r}
ggplot(covid_6)+
  geom_point(aes(x=date, y=deaths, color = stringency_index), position = 'jitter')+
  facet_wrap(~state)+
  theme(axis.text.x=element_text(angle=30))+
  labs(title='Cumulative Deaths Over Time at Varying Stringency Levels', x='Date Since Tracking Began', y='Cumulative Deaths')
```

---

## Comparing two numbers

```{r}
ggplot(data = covid) + geom_point(mapping = aes(x = tests, y = confirmed, color = state))
```

---


```{r}
ggplot(covid_6, aes(x=people_fully_vaccinated/population, y=deaths/confirmed, color=state))+geom_smooth()
```

---

```{r}
ggplot(covid_6) + geom_bar(aes(x=people_fully_vaccinated, y=deaths, color=state), stat='identity') + facet_wrap(~state)
```


---


```{r}
covid %>%
  mutate(avg_index = (stringency_index + government_response_index + containment_health_index + economic_support_index)/4) %>%
  ggplot() +
  geom_point(mapping = aes(x = avg_index, y = deaths))
```


I wanted to tidy up the data, I'm not sure how much significance I can derive from this plot considering that it looks a little confusing to me.
There are too many states, I hope I can show those states clearly



---

## Discrete x- or y-axis

```{r}
ggplot(covid)+geom_point(aes(y=state,x=tests)) +
  labs(title="US states v/s Number of tests")
```

I did not know how to change the scale of the x axis


---

```{r}
covid %>%
  filter(date == '2022-01-25') %>%
  ggplot(aes(x = reorder(state, -stringency_index), y = stringency_index)) + 
  geom_col() + 
  theme(axis.text.x = element_text(angle = 90)) +
  xlab('State') + ylab('Stringency Index') +
  ggtitle('State Stringency Indexes as of Jan 25th 2022') +
  geom_hline(yintercept = 30.6, color = 'red')

```


Couldn't add the horizontal line with the mean automatically (I had to separately calculate the mean and then manually input it into geom_hline). I wanted to plot a second layer of geom_point with the number of confirmed cases (to see how stringency index and covid cases were related), but it would just show the second layer (geom_point) without the original bar chart.


---

```{r}
ggplot(covid) + geom_bar(aes(x = state, fill = deaths))
```

---

```{r}
ggplot(covid) + geom_bar(aes(x = state, fill = deaths)) + coord_polar(theta = "x")
```

---

```{r}
ggplot(covid, mapping = aes(x = state, y = tests, color = state)) + 
  geom_point() +
  geom_boxplot() +
  labs(x = "Rate of Tests", y = "States", color = "States") +
  coord_flip()
```


I wanted to be able to use coord_polar() to see what this data would look like in that sort of graph, but the labels of each state made the visual of the graph itself less helpful.

---

```{r}
ggplot(covid)+geom_jitter(aes(y=state,x=tests, col=state)) +
  labs(title="US states v/s Number of tests")
```

---

```{r}
cov_data <- data.frame(x = covid_6$h6_facial__coverings, y = covid_6$confirmed)
plot_data <- data.frame(x = cov_data$x, y = log(cov_data$y))
X <- cov_data$x
Y <- log(cov_data$y)
indicator <- X
plot <- ggplot(plot_data) + geom_boxplot(varwidth = TRUE, aes(x = X, y = Y, colour = indicator)) 
plot  + ggtitle("Evaluating number of confirmed cases in relation to mask mandates") + xlab("Mask mandate (ordinal scale)") + ylab("Log of number of confirmed cases")
```



I wanted to do a scatterplot but I quickly noticed that with the way the data was (for facial covering) it would not work. I guess I just feel like I did not know what else to do in terms of plots and just do not know many of the plot functions available.

## Additional questions

Finally, I want to change the font and color of the title but keep it default, since I cannot find the code for font and the specific color.


I wanted to try making a pie chart, but I thought maybe the type of data we have wouldn't fit well as a pie chart (and every time I tried I got back very confusing plots)


I want to emphasize specific points and see what difference two states have.

---

# Coming Soon




```{r}
cleaned <- separate(covid_6, "date", c("Year", "Month", "Day"), sep = "-")
x <- cleaned %>% filter(Year == "2020" & Month <= "06")
ggplot(data = x) + 
    geom_point(mapping = aes(x = Month, y = confirmed, color = Month)) + 
    facet_grid(.~state) + 
    labs(title = "Confirmed cases in 6 major states from Jan 2020 to Jun 2020")
```


1. I wanted to add a best fit line for each state to show the tendency of the growth of confirmed cases in each state, but I don't know how.
2. I also wanted to change the scale of y-axis, so that it would be easier for visualization, but I weren't able to  change.


---


```{r}
covid %>%
  filter(date >= "2021-01-01", date <= "2022-01-01") %>%
  #filter(confirmed >= 1000, confirmed <= 50000) %>%
  filter(tests > mean(tests, na.rm = TRUE))%>%
  filter(vaccines > mean(vaccines, na.rm = TRUE))%>%
  ggplot(aes(x = date, y = tests, colour = state)) +
  geom_point(size = 4, alpha = 0.5, position = "jitter", shape = 19) +
  geom_point(aes(x = date , y = vaccines),size = 2, position = "jitter", shape = 17) +
  #geom_smooth()+
  scale_y_continuous(
    breaks = seq(from = 0, to = 1500000000, by = 5000000)
  , labels = seq(from = 0, , to = 1500000000, by = 5000000)
  )+
  scale_x_date(breaks = scales::breaks_width("1 month"), labels = scales::label_date_short())+
  ggtitle("Number of Tests vs Vaccines in 2021") + 
  theme(plot.title = element_text(color = "black", size = 15, face = "bold",hjust = 0.5, vjust = -5))+
  labs(x="", y="Number of Tests")
```

---

```{r}
ggplot(filter(weekly_6, !is.na(confirmed), !is.na(c6_stay_at_home_requirements))) +
    geom_line(aes(x=date, y = confirmed, color=state)) +
    facet_wrap(~ as.factor(c6_stay_at_home_requirements)) +
    scale_x_date(date_labels="%m/%y")
```

I was unable to label the facet_wraps as stay at home requirement levels. 


---


I wanted to be able to start the graph closer to December 2020 (when people started getting vaccinated) as this would reduce unnecessary white space in the plot 

---

```{r}
ggplot(daily_6) +
  geom_col(aes(x = state,y = sum(tests,na.rm=T)), fill="steelblue") +
  ggtitle("Number of Tests Performed from 2020-01-26 to 2022-01-25 per State") +
  labs(x="State", y = "Total Number of Tests Performed") +
  theme_minimal()
```


I wanted to be able to filter for a specific date, so for example I could create a plot that showed the total amount of daily tests per state (for a specific date).


---

```{r}
South_Northeast = (filter(covid, state == "Florida" | state == "Georgia" | state == "Alabama" | state == "New York" | state == "Connecticut" | state == "Massachusetts"))
ggplot(South_Northeast) +
  geom_col(aes(x = state, y = deaths/population, color = state)) + 
  labs(x = "State", y = "Deaths/Population", title = "State Death Rates Based on State Population")
```


Order the states so it showed the South States together and the Northeast states together



---

```{r}
ggplot(covid_6, aes(x = format(date, "%m"), y = government_response_index)) +
  geom_point(aes(color = state), size = 0.5, position = "jitter") +
  labs(x = "month")
```


I tried to represent month with abbreviated month such as Jan, but the x axis turns to be abbreviated month in alphabetical order but not the original order.

---

```{r}
covid_plot <- covid %>% filter(population < 1e6 |population > 1e7) %>% filter(!is.na(confirmed))

ggplot(covid_plot) + geom_point(aes(y = confirmed/population * 100, x = date, color = population), size = 0.75) + labs(title = "Cumulative % Population With Confirmed COVID-19", subtitle = "In US States With Population Under 1 Million or Over 10 Million", x = "Date", y = "% Population", color = "Population") + scale_color_distiller(palette = "Spectral") + theme_bw()
```


I wanted to compare Washington, Oregon, California, New York, Massachusetts, and Vermont and see how the trends in different regions compared, but I don't know how to add a column to the filtered dataset that could add either "NW" or "NE" based on the state in that row so that I could color the lines by region.

---

```{r}
CA1 <- filter(covid_6, state=="California", tests>=1000000)
ggplot(CA1, aes(x = date, y = deaths)) + geom_point(aes(color = state), size = 2) + stat_smooth(aes(group = state), color='green', geom = "line")
```

I want to choose more than one state.



---

```{r}
need1 <- filter(daily, vaccines > 0, state =='California')

p1 <- ggplot(need1, aes(x = date)) + geom_line(aes(y = ((cumsum(people_vaccinated)/population)*1000), color = 'red')) + geom_line(aes(y = ((deaths)), color = 'blue'), position = "jitter")+ labs(color = NULL) + scale_y_continuous(name = "Newly Deaths", limits = c(NA, NA), sec.axis = sec_axis(trans = ~. *.1, name = "vaccination rate ( %)")) 
p1 + scale_color_discrete( breaks = c("blue", "red"), label=c("newly death", "vaccine")) + theme_bw() + labs (title = "Vaccination VS. Newly Deaths") + theme(plot.title = element_text(size =15, face = "bold" ))
```

I want to calculate the rate of change of death initially, which (death of date2- death of date1)/death of date 1, but I don't know how to calculate it in ggplot. Therefore, I choose just put newly death.

---

```{r}
new_data <- filter(weekly, date>='2020-12-31', state %in% c("Massachusetts","New York","Connecticut","Pennsylvania")
                   ,c4_restrictions_on_gatherings!='NA')
ggplot(new_data,aes(x=date, y=confirmed)) + 
  geom_line(aes(color=state)) + 
  facet_grid(~as.factor(c4_restrictions_on_gatherings))
```



I want to split the date in year, month and date to simplify the x-axis of the graph.

I would like to be able to single out a single value by color. For example, in one of the NYT visualizations, there were multiple lines for each state, but the line for Massachusetts was green while all the others were grey.



# other


I wanted to create a facet grid following the instruction on Week 2 slides, but maybe because of the size of the data, they would run a very long time without actually providing me a plot.



I wish I knew how to adjust the labels to make the dates clearer, as well as the number of cases. It would have also been great if I was able to color coordinate the states on the plot and possibly add lines that follow the number of cases.

```{r}
y <- weekly_6$confirmed
x <- weekly_6$date
plot(x, y, main = "Confirmed Cases by Date",
  xlab = "Date",
  ylab = "Confirmed Cases")
```



# Data

I want to know how effective the vaccines are, but it seems like we do not have enough informations.

I want to include the confirmed cases, but it is difficult to scale confirmed cases, which it can fit into the data range of y axis. 




---

```{r}
ggplot(covid, aes(x = date, y = tests)) +
  geom_point(aes(color = state), size = 1) +
  stat_smooth(aes(group = state), color='black', alpha = 0.5, geom = "line")
```

smooth-en out the data points to fit the best-fit curves more
